An Introduction to Rocker: Docker Containers for R

### 2017-07-19 manuscript.Rmd
WIP

- **Portable**: from laptop to cloud
- **Interfaces**: Console-based (R, bash, etc) or RStudio\textsuperscript{\textregistered} Server, Shiny\textsuperscript{\textregistered} server
- **Sandboxed**: experiment without touching local system
- **Optimized**: community-tweaked configurations
- **Extensible**: interactively and persistent community extensions
- **Versioned**: reproducible environments

@@

@@

@@

Rocker images are designed to provide

Often Rocker images are not deployed directly, but used as a base image
onto which users add software for their own needs in a custom Dockerfile.
We offer a few recommendations here to streamline writing these extensions
and promote best practices.

- littler scripts

 To streamline the installation of R packages in downstream Dockerfiles,
 the `r-base` image provides `littler` package [@littler] and associated
 `install2.r` script, enabling users to install additional packages with
 a command such as:

    install2.r pkg1 pgk2 pkg3 ...

- Dockerfile best practices 
- automated builds, build triggers 
 - version-stable builds

>>>>

<!--  The main point here is that these are generic features that could be goals 
      of any effort to provide researchers with virtual computing environments.

      While most virtual computing environments (AMI images, scientific cloud
      platforms like cyverse, etc) provide some subset of these, they are most
      effective in concert.  -->

- **Portable**: from laptop to cloud
- **Interfaces**: Console-based (R, bash, etc) or RStudio\textsuperscript{\textregistered} Server, Shiny\textsuperscript{\textregistered} server
- **Sandboxed**: experiment without touching local system
- **Transparent**: Dockerfiles & automated builds provide a transparent and verifiable record of precisely what is installed.
- **Optimized**: community-tweaked configurations
- **Extensible**: interactively and persistent community extensions
- **Versioned**: reproducible environments

@@

Perhaps more importantly, Rocker containers are just as easy to run
on locally on most modern laptops on Windows, MacOS, and Linux-based
operating systems.

@@

### Transparent

Users can easily determine the software stack installed on any Rocker image
by examining the associated Dockerfile recipe, which provides a consise, human-readable
record of the installation.
All Rocker images use automated builds through DockerHub, which also acts as the central,
default repository distributing the images.  Using automated builds rather than uploading
pre-built image binaries to the Docker Hub avoids the potential for the build
not to match the recipe.  The corresponding Dockerfile is visible both on the Docker Hub
and in the linked GitHub repository of the Rocker project (<https://github.com/rocker-org), 
which provides a transparent versioned history of all changes made to these recipes,
as well as documentation, a community wiki, and issues trackers for discussing proposed
changes, bugs, improvements to the Dockerfiles and troubleshoot any issues users may encounter.

@@

<!-- Not intended as a guide to writing Dockerfiles, see http://ropenscilabs.github.io/r-docker-tutorial/05-dockerfiles.html -->

Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. To make these changes persistent, Rocker 
images can easily be extended by any user writing their own Dockerfiles based
on an appropriate Rocker image.

Rocker images provide a few utility functions to streamline this process, including the 
`littler` [@littler] scripts which provide a concise syntax for installing packages in Dockerfiles, e.g.

    RUN install2.r pkg1 pgk2 pkg3 ...

Users building on the version-tagged images will by default use the MRAN snapshot mirror
associated with the most recent date for which that image was current.  This ensures that
a Dockerfile building `FROM rocker/verse:3.4.0` will only install R package versions that were
available on CRAN on 2017-06-30, that is, the day R v3.4.1 was released.  (This default
can of course be overwritten by specifying a different CRAN mirror explicitly.)

Users writing their own Dockerfiles are encouraged to follow the same practices as the
Rocker Project, such as the [Dockerfile Best Practices](), the use of automated builds,
and when appropriate, versioned tags.  Users can ensure their automated builds are rebuilt
every time the relevant upstream DOckerfile is updated by using build triggers on Docker Hub.

Note that users can also preserve changes to Rocker images that they have modified interactively using
the `docker commit` mechanism, which creates a new binary image which can be pushed back to
a personal account on the Docker Hub.  While this is sometimes convenient, we encourage users to
consider writing Dockerfiles instead whenever possible, as this creates a more transparent
and reproducible mechanism to accomplish the same thing.

- Dockerfiles. community input.  [`hlapp/rpopgen`](https://github.com/NESCent/popgen-docker)
